Price of Anarchy for Mean Field Games

### 2018-07-04 PoA.tex
Minor point 10 fixed

This is a linear FBSDE of McKean-Vlasov type, which can be solved explicitly under mild assumptions (or at least in the case of time-independent coefficients which we will consider later. See Appendix \ref{appendix_solving_fbsdes}). Let $\bar{\eta}_t^{MFG}$, $\eta_t^{MFG}$, $\bar{x}_t^{MFG}$, and $v^{MFG}_t$ denote the solutions for this problem as described in the appendix so that $Y_t=\eta_t^{MFG}X_t+(\bar{\eta}_t^{MFG}-\eta_t^{MFG})\bar{x}_t^{MFG}$, $\mathbb{E}(Y_t)=\bar{\eta}_t^{MFG}\bar{x}_t^{MFG}$, $\mathbb{E}(X_t)=\bar{x}_t^{MFG}$, and $Var(X_t)=v^{MFG}_t$ provide a solution to the LQEMFG problem. Then from the appendix, we have:
\begin{equation}
\begin{split}
\dot{\bar{\eta}}^{MFG}_t+c^{MFG}(t)(b_2(t)+\bar{b}_2(t)) (\bar{\eta}^{MFG}_t)^2+(2b_1(t)+\bar{b}_1(t)) \bar{\eta}^{MFG}_t +q(t)+\bar{q}(t)(1-s(t))&=0,\\
\bar{\eta}^{MFG}_T-(q_T+\bar{q}_T(1-s_T))&=0,
\end{split}
\label{eq:eta_bar_MFG}
\end{equation}
\begin{equation}
\begin{split}
\dot{\eta}^{MFG}_t+a^{MFG}(t)b_2(t)(\eta^{MFG}_t)^2+2b_1(t)\eta^{MFG}_t+q(t)+\bar{q}(t)&=0, \\
\eta^{MFG}_T-(q_T+\bar{q}_T)&=0,
\end{split}
\end{equation}
\begin{equation}
\begin{split}
\dot{\bar{x}}^{MFG}_t&=(b_1(t)+\bar{b}_1(t)+c^{MFG}(t)(b_2(t)+\bar{b}_2(t))\bar{\eta}^{MFG}_t)\bar{x}^{MFG}_t, \\
\bar{x}^{MFG}_0&=\mathbb{E}(\xi).
\end{split}
\label{eq:x_bar_MFG}
\end{equation}
And thus,
\begin{equation*}
\bar{x}^{MFG}_t=\mathbb{E}(\xi) e^{\int_0^t(b_1(u)+\bar{b}_1(u)+c^{MFG}(u)(b_2(u)+\bar{b}_2(u))\bar{\eta}^{MFG}_u)du},
\end{equation*}
\begin{equation*}
v_t=Var(\xi)e^{\int_0^t 2(b_1(s)+a^{MFG}(s)b_2(s)\eta^{MFG}_s)ds}+\sigma^2 \int_0^t e^{2 \int_s^t (b_1(u)+a^{MFG}(u)b_2(u)\eta^{MFG}_u) du}ds.
\end{equation*}
Let $SC^{MFG}:=SC(\phi)$ for the feedback function specified by this solution, i.e.
\begin{equation*}
    \phi(t,x)=a^{MFG}(t)\eta_t^{MFG}x+\left(a^{MFG}(t)(\bar{\eta}_t^{MFG}-\eta_t^{MFG})+b^{MFG}(t)\bar{\eta}_t^{MFG} \right)\bar{x}_t^{MFG}.
\end{equation*}
Then we can compute the social cost as described in Section \ref{sub:MFG_formulation}:
\begin{equation*}
\begin{split}
    SC^{MFG}=&\frac{1}{2}[(q_T+\bar{q}_T)v_T^{MFG}+(q_T+\bar{q}_T(1-s_T)^2)(\bar{x}_T^{MFG})^2\\
    &+\int_0^T (q(t)+\bar{q}(t)+(r(t)+\bar{r}(t))(a^{MFG}(t)\eta_t^{MFG})^2)v_t^{MFG}\\
    &+(q(t)+\bar{q}(t)(1-s(t))^2+(r(t)+\bar{r}(t)(1-\bar{s}(t))^2)(c^{MFG}(t)\bar{\eta}_t^{MFG})^2)(\bar{x}_t^{MFG})^2dt],
\end{split}
\end{equation*}
where we have used the fact that:
\begin{equation*}
    \mathbb{E}(\phi(t,X_t))=c^{MFG}(t)\bar{\eta}_t^{MFG}\bar{x}_t^{MFG},
\end{equation*}
and:
\begin{equation*}
    Var(\phi(t,X_t))=(a^{MFG}(t)\eta_t^{MFG})^2v^{MFG}_t.
\end{equation*}

@@

\begin{proposition}
Assuming (\ref{eq:assumptions}), if furthermore,
\begin{equation*}
\begin{split}
    \left(s\bar{q}(s-1)+\bar{b}_1\bar{\eta}_t^{MKV}\right)\bar{x}_t^{MKV}&=0,\ \forall t \in [0,T] \\
    \bar{\eta}_t^{MKV}\bar{x}_t^{MKV}\left[(b^{MFG}-b^{MKV})b_2+(c^{MFG}-c^{MKV})\bar{b}_2\right]&=0,\ \forall t \in [0,T] \\
    s_T\bar{q}_T(s_T-1)\bar{x}_T^{MKV}&=0
\end{split}
\end{equation*}
then $PoA=1$.
\label{eq:lq_prop}
\end{proposition}

@@

\begin{proposition}
	Under assumption (\ref{eq:assumptions}), the difference in the social costs in the LQEMFG and LQEMKV problems can be represented by
	\begin{equation*}
		\Delta SC := SC^{MFG} - SC^{MKV} =  \frac{1}{2} \frac{(b_2 + \bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \int_0^T (\Delta \bar{\eta}_t \cdot \bar{x}_t^{MFG})^2 dt.
	\end{equation*}
\label{prop:diff_SC}
\end{proposition}
\begin{proof}
	The solutions $\bar{\eta}^{MFG}_t$ and $\bar{\eta}^{MKV}_t$ for the associated Riccati equations are well defined under assumption (\ref{eq:assumptions}). We notice that $\Delta\bar{\eta}_t$ defined in (\ref{eq:delta_eta}) satisfies the following first-order differential equation:
	\begin{equation*}
		\frac{d( \Delta{\bar{\eta}}_t)}{dt} = \gamma_t \Delta\bar{\eta}_t + \beta_t
		\qquad \Delta{\bar{\eta}}_T = \lambda \bar{\eta}_T^{MFG} - \bar{\eta}_T^{MKV}
	\end{equation*}
	with coefficients 
	\begin{equation*}
	\left\{
		\begin{array}{rcl}
			\gamma_t &=& \displaystyle -2 b_1 - 2\bar{b}_1 + \frac{(b_2 + \bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \left(\lambda \bar{\eta}_t^{MFG} + \bar{\eta}_t^{MKV} \right),\\
			\beta_t &=& \bar{b}_1 \lambda \bar{\eta}_t^{MFG} + (q + \bar{q}(1-s)^2) - \lambda (q + \bar{q}(1-s)).
		\end{array}
	\right.
	\end{equation*} 
	Since $q_T + \bar{q}_T(1-s_T)^2 = \bar{\eta}_T^{MKV}$ and $\lambda \bar{\eta}_0^{MFG} - \bar{\eta}_0^{MKV} = \Delta\bar{\eta}_0$, we deduce from equations (\ref{eq:SC_MFG_2}) and (\ref{eq:SC_MKV_2}) that:
	\begin{equation*}
		\begin{split}
			& SC^{MFG} - SC^{MKV}\\
			=& \frac{1}{2} \left[ \Delta\bar{\eta}_0 \bar{x}_0^2 - \Delta\bar{\eta}_T (\bar{x}_T^{MFG})^2 + \int_0^T \beta_t (\bar{x}_t^{MFG})^2 dt \right] \\
			=& \frac{1}{2} \int_0^T \left[ - \frac{d (\Delta\bar{\eta}_t (\bar{x}_t^{MFG})^2) }{dt}  + \left( 	\frac{d( \Delta{\bar{\eta}}_t)}{dt} - \gamma_t \Delta\bar{\eta}_t  \right)(\bar{x}_t^{MFG})^2 \right] dt\\
			=& \frac{1}{2}\int_0^T \left[- 2\Delta{\bar{\eta}}_t \bar{x}_t^{MFG} \dot{\bar{x}}_t^{MFG} - \gamma_t \Delta{\bar{\eta}}_t (\bar{x}_t^{MFG})^2 \right]dt \\
			=& \frac{1}{2} \int_0^T \Delta{\bar{\eta}}_t (\bar{x}_t^{MFG})^2 \left[-2(b_1 + \bar{b}_1 - \frac{(b_2+\bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \lambda \bar{\eta}_t^{MFG}) - \gamma_t\right] dt\\
			=& \frac{1}{2} \frac{(b_2 + \bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \int_0^T (\Delta \bar{\eta}_t \cdot \bar{x}_t^{MFG})^2 dt
		\end{split}
	\end{equation*}
	where we use the relationship $\dot{\bar{x}}_t = (a^x+a^{\bar{x}}+(a^y+a^{\bar{y}})\bar{\eta}_t)\bar{x}_t$ (see Appendix \ref{appendix_solving_fbsdes}) for the fourth equality.

@@

\begin{corollary}
    Assuming (\ref{eq:assumptions}), if the initial condition $\xi$ is such that $\mathbb{E}(\xi)=0$, then $PoA=1$.
\end{corollary}
\begin{proof}
    By the explicit equation for $\bar{x}_t$ (see Appendix \ref{appendix_solving_fbsdes}), $\bar{x}_0=0$ implies $\bar{x}_t=0,\ \forall t \in [0,T]$. Therefore by Proposition \ref{prop:diff_SC}, $\Delta SC=0$. From the explicit formula for the variance, $h_{var}>0$ and therefore by equation (\ref{eq:SC_MKV_2}), $SC^{MKV}>0$. Therefore, $PoA=1$.
\end{proof}

>>>>

This is a linear FBSDE of McKean-Vlasov type, which can be solved explicitly under mild assumptions (or at least in the case of time-independent coefficients which we will consider later. See Appendix \ref{appendix_solving_fbsdes}). Let $\bar{\eta}_t^{MFG}$, $\eta_t^{MFG}$, $\bar{x}_t^{MFG}$, and $v^{MFG}_t$ denote the solutions for this problem as described in the appendix so that $Y_t=\eta_t^{MFG}X_t+(\bar{\eta}_t^{MFG}-\eta_t^{MFG})\bar{x}_t^{MFG}$, $\mathbb{E}(Y_t)=\bar{\eta}_t^{MFG}\bar{x}_t^{MFG}$, $\mathbb{E}(X_t)=\bar{x}_t^{MFG}$, and $Var(X_t)=v^{MFG}_t$ provide a solution to the LQEMFG problem. Then from the appendix, we have:
\begin{equation}
\begin{split}
\dot{\bar{\eta}}^{MFG}_t+c^{MFG}(t)(b_2(t)+\bar{b}_2(t)) (\bar{\eta}^{MFG}_t)^2+(2b_1(t)+\bar{b}_1(t)) \bar{\eta}^{MFG}_t +q(t)+\bar{q}(t)(1-s(t))&=0,\\
\bar{\eta}^{MFG}_T-(q_T+\bar{q}_T(1-s_T))&=0,
\end{split}
\label{eq:eta_bar_MFG}
\end{equation}
\begin{equation}
\begin{split}
\dot{\eta}^{MFG}_t+a^{MFG}(t)b_2(t)(\eta^{MFG}_t)^2+2b_1(t)\eta^{MFG}_t+q(t)+\bar{q}(t)&=0, \\
\eta^{MFG}_T-(q_T+\bar{q}_T)&=0,
\end{split}
\end{equation}
\begin{equation}
\begin{split}
\dot{\bar{x}}^{MFG}_t&=(b_1(t)+\bar{b}_1(t)+c^{MFG}(t)(b_2(t)+\bar{b}_2(t))\bar{\eta}^{MFG}_t)\bar{x}^{MFG}_t, \\
\bar{x}^{MFG}_0&=\mathbb{E}(\xi).
\end{split}
\label{eq:x_bar_MFG}
\end{equation}
And thus,
\begin{equation}
\bar{x}^{MFG}_t=\mathbb{E}(\xi) e^{\int_0^t(b_1(u)+\bar{b}_1(u)+c^{MFG}(u)(b_2(u)+\bar{b}_2(u))\bar{\eta}^{MFG}_u)du},
\label{eq:x_bar_MFG_explicit}
\end{equation}
\begin{equation}
v_t=Var(\xi)e^{\int_0^t 2(b_1(s)+a^{MFG}(s)b_2(s)\eta^{MFG}_s)ds}+\sigma^2 \int_0^t e^{2 \int_s^t (b_1(u)+a^{MFG}(u)b_2(u)\eta^{MFG}_u) du}ds.
\label{eq:v_t}
\end{equation}
Let $SC^{MFG}:=SC(\phi)$ for the feedback function specified by this solution, i.e.
\begin{equation*}
    \phi(t,x)=a^{MFG}(t)\eta_t^{MFG}x+\left(a^{MFG}(t)(\bar{\eta}_t^{MFG}-\eta_t^{MFG})+b^{MFG}(t)\bar{\eta}_t^{MFG} \right)\bar{x}_t^{MFG}.
\end{equation*}
Then we can compute the social cost as described in Section \ref{sub:MFG_formulation}:
\begin{equation*}
\begin{split}
    SC^{MFG}=&\frac{1}{2}[(q_T+\bar{q}_T)v_T^{MFG}+(q_T+\bar{q}_T(1-s_T)^2)(\bar{x}_T^{MFG})^2\\
    &+\int_0^T (q(t)+\bar{q}(t)+(r(t)+\bar{r}(t))(a^{MFG}(t)\eta_t^{MFG})^2)v_t^{MFG}\\
    &+(q(t)+\bar{q}(t)(1-s(t))^2+(r(t)+\bar{r}(t)(1-\bar{s}(t))^2)(c^{MFG}(t)\bar{\eta}_t^{MFG})^2)(\bar{x}_t^{MFG})^2dt],
\end{split}
\end{equation*}
where we have used the fact that:
\begin{equation*}
    \mathbb{E}(\phi(t,X_t))=c^{MFG}(t)\bar{\eta}_t^{MFG}\bar{x}_t^{MFG},
\end{equation*}
and:
\begin{equation*}
    Var(\phi(t,X_t))=(a^{MFG}(t)\eta_t^{MFG})^2v^{MFG}_t.
\end{equation*}

@@

\begin{proposition}
Assuming (\ref{eq:assumptions}), if furthermore,
\begin{equation*}
\begin{split}
    \left(s\bar{q}(s-1)+\bar{b}_1\bar{\eta}_t^{MKV}\right)\bar{x}_t^{MKV}&=0,\ \forall t \in [0,T] \\
    \bar{\eta}_t^{MKV}\bar{x}_t^{MKV}\left[(b^{MFG}-b^{MKV})b_2+(c^{MFG}-c^{MKV})\bar{b}_2\right]&=0,\ \forall t \in [0,T] \\
    s_T\bar{q}_T(s_T-1)\bar{x}_T^{MKV}&=0,
\end{split}
\end{equation*}
then $PoA=1$.
\label{eq:lq_prop}
\end{proposition}

@@

\begin{proposition}
	Under assumption (\ref{eq:assumptions}), the difference in the social costs in the LQEMFG and LQEMKV problems can be represented by
	\begin{equation*}
		\Delta SC := SC^{MFG} - SC^{MKV} =  \frac{1}{2} \frac{(b_2 + \bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \int_0^T (\Delta \bar{\eta}_t \cdot \bar{x}_t^{MFG})^2 dt.
	\end{equation*}
\label{prop:diff_SC}
\end{proposition}
\begin{proof}
	The solutions $\bar{\eta}^{MFG}_t$ and $\bar{\eta}^{MKV}_t$ for the Riccati equations (\ref{eq:eta_bar_MFG}) and (\ref{eq:eta_bar_MKV}), respectively, are well defined under assumption (\ref{eq:assumptions}) (see Appendix \ref{appendix_solving_fbsdes}). We notice that $\Delta\bar{\eta}_t$ defined in (\ref{eq:delta_eta}) satisfies the following first-order differential equation:
	\begin{equation*}
		\frac{d( \Delta{\bar{\eta}}_t)}{dt} = \gamma_t \Delta\bar{\eta}_t + \beta_t,
		\qquad \Delta{\bar{\eta}}_T = \lambda \bar{\eta}_T^{MFG} - \bar{\eta}_T^{MKV}
	\end{equation*}
	with coefficients:
	\begin{equation*}
	\left\{
		\begin{array}{rcl}
			\gamma_t &=& \displaystyle -2 b_1 - 2\bar{b}_1 + \frac{(b_2 + \bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \left(\lambda \bar{\eta}_t^{MFG} + \bar{\eta}_t^{MKV} \right),\\
			\beta_t &=& \bar{b}_1 \lambda \bar{\eta}_t^{MFG} + (q + \bar{q}(1-s)^2) - \lambda (q + \bar{q}(1-s)).
		\end{array}
	\right.
	\end{equation*} 
	Since $q_T + \bar{q}_T(1-s_T) = \bar{\eta}_T^{MFG}$, $q_T + \bar{q}_T(1-s_T)^2 = \bar{\eta}_T^{MKV}$ and $\lambda \bar{\eta}_0^{MFG} - \bar{\eta}_0^{MKV} = \Delta\bar{\eta}_0$, we deduce from equations (\ref{eq:SC_MFG_2}) and (\ref{eq:SC_MKV_2}) that:
	\begin{equation*}
		\begin{split}
			& SC^{MFG} - SC^{MKV}\\
			=& \frac{1}{2} \left[ \Delta\bar{\eta}_0 \bar{x}_0^2 - \Delta\bar{\eta}_T (\bar{x}_T^{MFG})^2 + \int_0^T \beta_t (\bar{x}_t^{MFG})^2 dt \right] \\
			=& \frac{1}{2} \int_0^T \left[ - \frac{d (\Delta\bar{\eta}_t (\bar{x}_t^{MFG})^2) }{dt}  + \left( 	\frac{d( \Delta{\bar{\eta}}_t)}{dt} - \gamma_t \Delta\bar{\eta}_t  \right)(\bar{x}_t^{MFG})^2 \right] dt\\
			=& \frac{1}{2}\int_0^T \left[- 2\Delta{\bar{\eta}}_t \bar{x}_t^{MFG} \dot{\bar{x}}_t^{MFG} - \gamma_t \Delta{\bar{\eta}}_t (\bar{x}_t^{MFG})^2 \right]dt \\
			=& \frac{1}{2} \int_0^T \Delta{\bar{\eta}}_t (\bar{x}_t^{MFG})^2 \left[-2\left(b_1 + \bar{b}_1 - \frac{(b_2+\bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \lambda \bar{\eta}_t^{MFG}\right) - \gamma_t\right] dt\\
			=& \frac{1}{2} \frac{(b_2 + \bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2} \int_0^T (\Delta \bar{\eta}_t \cdot \bar{x}_t^{MFG})^2 dt,
		\end{split}
	\end{equation*}
	where we use equation (\ref{eq:x_bar_MFG}) for the fourth equality.

@@

\begin{corollary}
    Assuming (\ref{eq:assumptions}), if the initial condition $\xi$ is such that $\mathbb{E}(\xi)=0$, then $PoA=1$.
\end{corollary}
\begin{proof}
    By equation (\ref{eq:x_bar_MFG_explicit}), $\mathbb{E}(\xi)=0$ implies $\bar{x}^{MFG}_t=0,\ \forall t \in [0,T]$. Therefore by Proposition \ref{prop:diff_SC}, $\Delta SC=0$. From equation (\ref{eq:v_t}), $v_t>0$ which implies $h_{var}>0$. Therefore by equation (\ref{eq:SC_MKV_2}), $SC^{MKV}>0$. We conclude that $PoA=1$.
\end{proof}
